Sequence Specific Modeling of \emph{E.~coli} Cell-Free Protein Synthesis

### 2017-05-19 Paper_v1.tex
Fixed issues w/the refs

\section{Introduction}
Cell-free protein expression has become a widely used research tool in systems and synthetic biology, and a promising technology for personalized protein production.
Cell-free systems offer many advantages for the study, manipulation and modeling of metabolism compared to \textit{in vivo} processes.
Central amongst these is direct access to metabolites and the biosynthetic machinery without the interference of a cell wall or the complications associated with cell growth.
This allows interrogation of the chemical environment while the biosynthetic machinery is operating, potentially at a fine time resolution.
Cell-free protein synthesis (CFPS) systems are arguably the most prominent examples of cell-free systems used today \cite{Jewett:2008aa}.
However, CFPS is not new; CFPS in crude \textit{E.~coli} extracts has been used since the 1960s to explore fundamental biological mechanisms \cite{MATTHAEI:1961aa,NIRENBERG:1961aa}.
Today, cell-free systems are used in a variety of applications ranging from therapeutic protein production \cite{Lu:2014aa} to synthetic biology \cite{Hodgman:2012aa}.
However, if CFPS is to become a mainstream technology for advanced applications such as point of care manufacturing \cite{Pardee2016248}, we must first understand the performance limits and costs of these systems \cite{Jewett:2008aa}.
One tool to address these questions is constraint based modeling.

@@

\subsubsection{Potential alternative metabolic optima}
Optimal flux distributions predicted using constraint based approaches may not always be unique.
Alternative optimal solutions have the same objective value, e.g., productivity, but different metabolic flux distributions.
Several techniques such as flux variability analysis (FVA) \cite{Mahadevan2003264,Schuetz119} or mixed-integer approaches \cite{LEE2000711} can estimate alternative optima.
In this study, we used group knockout analysis to estimate potential alternative optimal solutions for CAT production constrained by experimental measurements (Fig.~\ref{fig:norm}).
In this approach, groups of reactions were removed from the metabolic network, and the translation rate was maximized.
The difference in the productivity and flux distribution of the nominal and altered system was then calculated.
Knockout analysis identified pathways required for CAT production; for example, deletion of the glycolysis/gluconoeogenesis or oxidative phosphorylation pathways resulted in no CAT production.
Likewise, there were pathway knockouts that had no effect on productivity or the metabolic flux distribution, such as removal of isoleucine, leucine, histidine and valine biosynthesis.
Globally, the constraint based simulation reached the same optimal CAT productivity for 40\% of the pairwise knockouts, while 92\% of these solutions had different flux distributions compared with the wild-type.
For example, one of the features of the predicted optimal metabolic flux distribution was a high flux through the Entner-Douodoroff (ED) pathway.
Removal of the ED pathway had no effect on the CAT productivity compared to the absence of knockouts (Fig.~\ref{fig:norm}A).
Pairwise knockouts of the ED pathway and other subgroups (i.e. pentose phosphate pathway, cofactors, folate metabolism, etc.) also resulted in the same optimal CAT productivity.
However, there was a difference in the flux distribution with these knockouts (Fig.~\ref{fig:norm}B); thus, alternative optimal metabolic flux distributions exist for CAT production, despite experimental constraints.
In addition, knockouts of amino acid biosynthesis reactions had no effect on the productivity with the exception of alanine, aspartate, asparagine, glutamate and glutamine biosynthesis reactions, since amino acids were available in the media.
Ultimately, to determine the metabolic flux distribution occurring in CFPS, we need to add additional constraints to the flux estimation calculation.
For example, thermodynamic feasibility constraints may result in a better depiction of the flux distribution \cite{Henry:2007,Hamilton:2013}, and $^{13}$C labeling in CFPS could provide significant insight.
However, while $^{13}$C labeling techniques are well established for \emph{in~vivo} processes \cite{Zamboni:2009}, application of these techniques to CFPS remains an active area of research.

>>>>

\section{Introduction}
Cell-free protein expression has become a widely used research tool in systems and synthetic biology, and a promising technology for personalized protein production.
Cell-free systems offer many advantages for the study, manipulation and modeling of metabolism compared to \textit{in vivo} processes.
Central amongst these is direct access to metabolites and the biosynthetic machinery without the interference of a cell wall or the complications associated with cell growth.
This allows interrogation of the chemical environment while the biosynthetic machinery is operating, potentially at a fine time resolution.
Cell-free protein synthesis (CFPS) systems are arguably the most prominent examples of cell-free systems used today \cite{Jewett:2008aa}.
However, CFPS is not new; CFPS in crude \textit{E.~coli} extracts has been used since the 1960s to explore fundamental biological mechanisms \cite{MATTHAEI:1961aa,NIRENBERG:1961aa}.
Today, cell-free systems are used in a variety of applications ranging from therapeutic protein production \cite{Lu:2014aa} to synthetic biology \cite{Hodgman:2012aa}.
However, if CFPS is to become a mainstream technology for advanced applications such as point of care manufacturing \cite{Pardee:2016aa}, we must first understand the performance limits and costs of these systems \cite{Jewett:2008aa}.
One tool to address these questions is constraint based modeling.

@@

\subsubsection{Potential alternative metabolic optima}
Optimal flux distributions predicted using constraint based approaches may not always be unique.
Alternative optimal solutions have the same objective value, e.g., productivity, but different metabolic flux distributions.
Several techniques such as flux variability analysis (FVA) \cite{Mahadevan2003264,Schuetz:2007aa} or mixed-integer approaches \cite{LEE2000711} can estimate alternative optima.
In this study, we used group knockout analysis to estimate potential alternative optimal solutions for CAT production constrained by experimental measurements (Fig.~\ref{fig:norm}).
In this approach, groups of reactions were removed from the metabolic network, and the translation rate was maximized.
The difference in the productivity and flux distribution of the nominal and altered system was then calculated.
Knockout analysis identified pathways required for CAT production; for example, deletion of the glycolysis/gluconoeogenesis or oxidative phosphorylation pathways resulted in no CAT production.
Likewise, there were pathway knockouts that had no effect on productivity or the metabolic flux distribution, such as removal of isoleucine, leucine, histidine and valine biosynthesis.
Globally, the constraint based simulation reached the same optimal CAT productivity for 40\% of the pairwise knockouts, while 92\% of these solutions had different flux distributions compared with the wild-type.
For example, one of the features of the predicted optimal metabolic flux distribution was a high flux through the Entner-Douodoroff (ED) pathway.
Removal of the ED pathway had no effect on the CAT productivity compared to the absence of knockouts (Fig.~\ref{fig:norm}A).
Pairwise knockouts of the ED pathway and other subgroups (i.e. pentose phosphate pathway, cofactors, folate metabolism, etc.) also resulted in the same optimal CAT productivity.
However, there was a difference in the flux distribution with these knockouts (Fig.~\ref{fig:norm}B); thus, alternative optimal metabolic flux distributions exist for CAT production, despite experimental constraints.
In addition, knockouts of amino acid biosynthesis reactions had no effect on the productivity with the exception of alanine, aspartate, asparagine, glutamate and glutamine biosynthesis reactions, since amino acids were available in the media.
Ultimately, to determine the metabolic flux distribution occurring in CFPS, we need to add additional constraints to the flux estimation calculation.
For example, thermodynamic feasibility constraints may result in a better depiction of the flux distribution \cite{Henry:2007,Hamilton:2013}, and $^{13}$C labeling in CFPS could provide significant insight.
However, while $^{13}$C labeling techniques are well established for \emph{in~vivo} processes \cite{Zamboni:2009}, application of these techniques to CFPS remains an active area of research.
